A formal approach but practical approach in train regulation

### 2025-06-18 paper/main.tex
le régulateur est maintenant un triplet

Routes are not computed in advance. The schedule is renewed every "day"\footnote{We use a time scale factor: one of our "day" corresponds to 20 minutes in real time.}  and we introduce some randomness in the process: beyond some routine daily planning, some trains are inserted/deleted, some trains departure/destination may change, and so on. We simulate also delays. Finally, in our model, we distinguish freight and  passengers routes giving priority to the latter. Routes are recomputed each time a train has finished its duty. The general motivation of this paper is to propose a lightweight scheduling solution that is adapted to the fully automatic control, that takes into account the distributed nature of the system, that is reliable and does not suffer from interlocking.

@@

The regulator is in charge of the execution of the planning. The regulator uses some tokens to ensure a reliable circulation. For each critical resource (either a block or a switch), we associate a token level with each train involved.

More technically, the regulator is characterized by 5 fields. First, it has a general program, next mentioned as $E$. The program is the plan of the regulator. The regulator will follow its rules. A program is a sequence of statements. Each statement is composed of a guard and some actions. The guard has the shape: $ev(tid, n)$ meaning that train $tid$ just entered in the $n$-th block of its route. In case such an event occurs, the associated actions are performed. Here is the grammar of the regulator program:
%All statements are guarded by a condition $ev(tid,n)$ stating that train $tid$ reached the $n$-th step of its route. The atoms update the regulator fields. 
%A program is given as a sequence of statements. Those follow the grammar:

@@

The waiting field is a function $W: \blocks \times \Nat \to \trains_0$. It states that block must be protected. More precisely, suppose that for some block $pid$, some train $tid$ tries to enter the block. If $W(pid,T(pid)) = tid$, then, the train shall wait the token is released. The value $W(pid, n) = 0$ means the train did not reach the current position $pid$.

The history field $H: \trains \to \blocks$ gives the last position of each train recorded  by the regulator. \todo{A vérifier.}

The switch field is a function $S: \turnouts \to \internalState$ giving the network configuration. Again, the atom $\turn(sid,s)$ updates the internal state of $id$: $S \cdot \turn(sid,s) \triangleq S[sid \leftarrow s]$.

So, the regulator is described as a tuple $R = (E, T, S, W, H)$. Applying $\incr(j)$ to the regulator is applying it to the token field: $R \cdot \incr(j) = (E, T \cdot \incr(j), S, W, B, H)$. We use the same convention for the other atoms.

>>>>

Routes are computed on the fly. The schedule is renewed every "day"\footnote{We use a time scale factor: one of our "day" corresponds to 20 minutes in real time.}  and we introduce some randomness in the process: beyond some routine daily planning, some trains are inserted/deleted, some trains departure/destination may change, and so on. We simulate also delays. Finally, in our model, we distinguish freight and  passengers routes giving priority to the latter. Routes are recomputed each time a train has finished its duty. The general motivation of this paper is to propose a lightweight scheduling solution that is adapted to the fully automatic control, that takes into account the distributed nature of the system, that is reliable and does not suffer from interlocking.

@@

The regulator is in charge of the execution of the planning. For that sake, we use tokens. For each critical resource (either a block or a switch), we associate a token level with each train involved.

More technically, the regulator is characterized by 3 fields. First, it has a general program, next mentioned as $E$. The program is the plan of the regulator. The regulator will follow its rules. A program is a sequence of statements, each statement being composed of a guard and some actions. The guard has the shape: $ev(tid, n)$ meaning that train $tid$ just entered in the $n$-th block of its route. In case such an event occurs, the associated actions are performed. Here is the grammar of the regulator program:
%All statements are guarded by a condition $ev(tid,n)$ stating that train $tid$ reached the $n$-th step of its route. The atoms update the regulator fields. 
%A program is given as a sequence of statements. Those follow the grammar:

@@

The waiting field is a set $W \subseteq \blocks \times \Nat \times \trains$. A triple $(pid, n, tid) \in W$  meaning intention is that block $pid$ must be protected against train $tid$ entrance as long as $T(pid) < n$.

So, the regulator is described as a tuple $R = (E, T, W)$. Applying $\incr(pid)$ to the regulator is applying it to the token field: $R \cdot \incr(pid) = (E, T \cdot \incr(pid), W)$. We use the same convention for the other atoms.

Finally, the network configuration is given as  $\sigma: \turnouts \to \internalState$. Again, the atom $\turn(sid,s)$ updates the internal state of $sid$: $\sigma \cdot \turn(sid,s) \triangleq \sigma[sid \leftarrow s]$.
